--- 
title: "Operationalizing a national protocol for digital point counts"
author: "Alexander G. MacPhail"
date: "`r format(Sys.time(), '%d %B, %Y')`"
site: bookdown::bookdown_site
output: bookdown::bs4_book
url: https:/abbiodiversity.github.io/ndpc
description: |
  Digital point counts
---

```{r include=FALSE, eval=T}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
knitr::is_html_output()

library(tidyverse)
library(lubridate)
library(fs)
library(sf)
library(gganimate)

```

# Welcome

Welcome to the protocol for conducting a **digital avian point count** using an autonomous recording unit or handheld recording device. In collaboration with partners from [Environment and Climate Change Canada](), the [Breeding Bird Survey](), the [Ontario Breeding Bird Atlas](), the [Bioacoustic Unit](), the [Alberta Biodiversity Monitoring Institute]() and the [Boreal Avian Modeling Project](), this user-friendly protocol aims to establish a standard method for collecting data on avian populations in a digital format, making it accessible and easy for citizen scientists to collect and contribute to avian research and population monitoring.

```{r, eval=T, include=F, warning=F, message=F} 

locs <- dir_ls(".", recurse = T, regexp = "Locations") %>%
  map_dfr(~read_csv(.)) %>%
  select(location, latitude, longitude)

pcs <- dir_ls(".", recurse = T, regexp = "Tasks") %>%
  map_dfr(~read_csv(.)) %>%
  select(location, recordingDate) %>%
  inner_join(., locs, by = c("location" = "location")) %>%
  filter(!is.na(latitude)) %>%
  mutate(year = year(recordingDate)) %>%
  sample_frac(0.03) %>%
  st_as_sf(coords = c("longitude","latitude")) %>%
  st_make_valid() %>%
  st_set_crs(4269)

ggplot() +
  coord_sf(crs = 4269) +
  geom_sf(pcs, mapping = aes(colour = year)) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_colour_viridis_c() +
  transition_states(recordingDate, state_length = 0,
                    transition_length = c(rep(1, 10), 9, rep(1, 9))) + 
  shadow_mark() +
  ggtitle('Date of Survey: {closest_state}',
          subtitle = "Count: {frame}")
```

- Use of autonomous recording units (ARUs) and handheld recording devices to conduct avian point counts, in collaboration with leading avian research organizations.
- Standardized protocol for data collection to ensure consistency and comparability of results, following established guidelines
- Digital format for data collection, making it easily accessible and shareable among researchers from various organizations.
- Citizen scientists can participate in avian research and contribute to scientific knowledge, working in partnership with established organizations
- No prior experience or expertise required, making it user-friendly for all skill levels, with support and guidance from expert stakeholders

## Frequently Asked Questions {.unnumbered}

- [What recorder can I use?]()
- [Where do I submit my recordings?]()





